---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# scoreMVAQ

<!-- badges: start -->
<!-- badges: end -->

The goal of scoreMVAQ is to convert the 14 responses from the Michigan Vision-related Anxiety Questionnaire (MVAQ) to two domain scores: Anxiety due to Rod Function (ARF) and Anxiety due to Cone Function (ACF)

## Installation

You can install the development version of scoreMVAQ from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("chrisandrewsphd/scoreMVAQ")
```

## Examples

### Example 1

This is a basic example which shows you how to score data from a single individual. `justme` is a data.frame with 15 variables and 1 row.  The 15 variables are the patient identifier followed by their 14 responses.

```{r example1}
library(scoreMVAQ)

justme <- read.table(
  text = "Me,0,0,0,0,0,0,0,0,1,1,1,0,0,1",
  sep = ",")
justme

scoreMVAQ(justme)
```

### Example 2

If you need to score many respondents (or the same respondent at several visits), you can include them all in a single data.frame and call `scoreMVAQ()`. The data.frame must have 15 columns still but can have more than 1 row.

```{r example2}
library(scoreMVAQ)

justme <- read.table(
  text = "Me,0,0,0,0,0,0,0,0,1,1,1,0,0,1",
  sep = ",")
justyou <- data.frame(ID="You", matrix(nrow=1, sample(0:3, 14, TRUE)))
andyoutoo <- data.frame(ID="You 2", matrix(nrow=1, sample(0:3, 14, TRUE)))

names(justme) <- names(justyou) <- names(andyoutoo) <- 
  c("ID", sprintf("Q%0.2d", seq(14)))

justus <- rbind(justme, justyou, andyoutoo)
justus

scoreMVAQ(justus, verbose = 1)
```

The option `verbose = 1` provides some messaging during the execution of `scoreMVAQ()`.
