---
title: "scoreMVAQ"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{scoreMVAQ}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>")

library(scoreMVAQ)
```

# The Michigan Vision-related Anxiety Questionnaire

The MVAQ (https://www.ncbi.nlm.nih.gov/pmc/articles/PMC8184579/) is a 14-item questionnaire to assess anxiety related to visual function among patients with inherited retinal degeneration. A patient's responses are used to quantify visual function in 2 domains:

* Anxiety Related to Rod Function (ARF) and
* Anxiety Related to Cone Function (ACF).

The MVAQ itself is copyrighted material of the University of Michigan. All rights reserved. For licensing information, contact the University of Michigan Technology Transfer Office at `techtransfer@umich.edu`. This scoring software is openly available.

# Scoring the MVAQ

A graded response model (GRM) was built from 128 original patients at the University of Michigan. The R package mirt was used to fit the GRM and can be used to score responses from new patients. The R package scoreMVAQ contains the (deidentified) mirt model objects and a wrapper function `scoreMVAQ()` that uses 14 responses to generate 2 domain scores (and standard errors).

## Installation

The scoreMVAQ package is available from GitHub. An easy way to install scoreMVAQ is 

```{r installelaborate, echo = FALSE}
# if (!isTRUE(require(scoreMVAQ))) {
#   if (!isTRUE(require("devtools"))) install.packages("devtools")
#   devtools::install_github("ChrisAndrewsPhd/scoreMVAQ")
# }
```

```{r install}
# install.packages("devtools")
# devtools::install_github("ChrisAndrewsPhd/scoreMVAQ")
```

Installation is required the first time scoreMVAQ is used on a machine, after a major version change of R (e.g. 4.2.1 to 4.3.0), or to obtain the most recent version of the package.

## Examples from Package

The scoreMVAQ package has a few simple examples in the helpfile for `scoreMVAQ()`

```{r packageexamples}
library(scoreMVAQ)

# ?scoreMVAQ # open help file

example(scoreMVAQ) # run examples in help file
```


## A Typical Use Case

The responses of an administration of the MVAQ should be stored in a single row of a dataset. For example, in a cross-sectional study, each patient has one row and 14 of the columns are the responses to the 14 items.  In a longitudinal study, each patient visit has a row and 14 of the columns are the responses. If the only columns in the data.frame are an identifier and the 14 responses *in order*, then the function `scoreMVAQ()` can be used with the dataset directly.  However, it is often the case that the data.frame contains additional columns that should not be passed to `scoreMVAQ()`, such as VisitNumber, Sex, ExamDate, ....

### Data Preparation

Read the data from a file and create a data.frame with 15 columns. The first column is not used. It is carried along.  `scoreMVAQ()` does not reorder the rows. A generic approach could just use, e.g., seq(nrow(patientresponses)). In this example, The PatID and VisitID variables are pasted to create the identifier

```{r dataprep}
patientresponses <- read.csv(
  system.file("extdata/sampleMVAQ.csv", package = "scoreMVAQ", mustWork = TRUE))


dim(patientresponses) # has 61 columns (need 60)
patientresponses # Columns for PatID and VisitID

dat_for_MVAQ <- data.frame(
  ID_Visit = paste(
    patientresponses$PatID,
    patientresponses$VisitID,
    sep = "|"),
  patientresponses[, seq(from = 3, by = 1, length = 14)])
dim(dat_for_MVAQ)
dat_for_MVAQ
```

### Scoring

The function `scoreMVAQ()` returns a list with 2 elements.  Each element is a data.frame with 3 columns: ID and the 2 domain scores. Each data.frame has the same number of rows as the input.

```{r applyscoreMVAQ}
MVAQlist <- scoreMVAQ::scoreMVAQ(dat_for_MVAQ, verbose = 0)
str(MVAQlist) # list with 2 elements

MVAQthetas <- MVAQlist[["thetas"]] # domain estimates
MVAQses <- MVAQlist[["ses"]] # standard error estimates
```

### Wrap up

The `scoreMVAQ()` output can be saved as an R object, written to an external file, used to create summary data, plotted, ...

```{r postprocessing}
patientthetas <- data.frame(patientresponses, MVAQthetas[, -1])
patientthetas

# save(
#   MVAQlist, dat_for_MVAQ, patientthetas,
#   file = "./example.RData")
# 
# write.csv(
#   MVAQthetas,
#   file = "./examplethetas.csv",
#   row.names = FALSE, na = "")
```

```{r table, eval = requireNamespace("kableExtra") & requireNamespace("knitr")}
library(knitr)
library(kableExtra)

kableExtra::kable_styling(
  knitr::kable(
    sapply(
      subset(
        patientthetas,
        subset = VisitID == "baseline",
        select = names(MVAQthetas[, -1])),
      quantile, na.rm = TRUE),
    digits = 2,
    caption = "Summary of Baseline ARF and ACF"),
  full_width = FALSE)
```

```{r plot, fig.height = 4, fig.width = 6}
repeater <- subset(
  patientthetas,
  subset = PatID == "Repeater",
  select = c("VisitID", "ARF", "ACF"))

par(mar = c(4.1, 4.1, 3, 1), las = 1)
matplot(
  repeater[, c("ARF", "ACF")],
  type = "b", pch = 16, lty = 1, lwd = 2, col = c("green", "grey"),
  ylab = "Anxiety", ylim = c(-1.5, 2.6),
  xlab = "Visit", xaxt = "n",
  main = "Longitudinal Scores for Patient 'Repeater'")
axis(
  side = 1,
  at = seq(nrow(repeater)),
  labels = repeater$VisitID)
legend(
  "top", bty = "n", col = c("green", "grey"), lwd = 2, pch = 16,
  legend = c("ARF", "ACF"), ncol = 2)
```
